---
title: "Change in Alochol Consumption and Current Vacination Status in Sweden"
description: |
  A short description of the post.
categories:
  - COVID-19
author:
  - name: Jaffie Yang
    url: https://etc5523-2021.github.io/blog-myan0065/about.html
    affiliation: Monash NUMBATs
    affiliation_url: https://numbat.space/
date: 09-12-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      out.width = "80%",
                      fig.align = "center")
```

```{r package}
library(readxl)
library(here)
library(ggplot2)
library(gghighlight)
library(treemapify)
library(patchwork)
library(dplyr)
library(tidyverse)
library(tidytext)
library(lubridate)
library(naniar)
library(kableExtra)
library(plotly)
library(DT)
library(scales)
```

```{r read-data}
vacc_se <- read_csv(here("data/data_se.csv"))
world <- read_csv(here("data/owid-covid-data.csv"))
surv_se <- read_csv(here("data/wdata_COVID_2207.csv"))
surv_dict <- read_excel(here("data/Codebook_2207.xlsx"), sheet = 2, range = "b1:d40")
```

## Introduction 

In this post, we will continue to explore the COVID status in Europe, but in a different country, Sweden. Two key topics will be addressed, change in alcohol consumption during pandemic and vaccination status in Sweden.

## Data Description 📄

### COVID -19 Survey Data

This survey data is obtained from <a href=https://covid19dataportal.se/data_types/all/data/> Swedish COVID-19 Data Portal</a>, it is publicly accessible and under MIT license. It is a survey done by 40,060 people from `r length(unique(surv_se$country_label))` different European countries, all the entries were done between April to July in 2020 when COVID-19 hit Europe. There are `r ncol(surv_se)` variables and `r nrow(surv_se)` observations, data dictionary and missing values are shown in Figure \@ref(fig:surv-dict) and Figure \@ref(fig:surv-miss).

```{r surv-dict, fig.cap = "COVID-19 Survey 2020, Europe"}
surv_dict %>% 
  datatable(options = list(pageLength = 5,
                            columnDefs = list(list(className = 'dt-left', targets = 1:3))),
            colnames = c("Variable", "Original Question / Description", "Value Labels"),
            caption = "Data Dictionary")
```
```{r surv-miss, fig.cap = "Summary of Missing Value in COVID-19 Survey Data"}
# surv_se %>% 
#   filter(country_label == "Sweden") %>% 
# miss_var_summary() %>% 
#     datatable(options = list(pageLength = 5,
#                             columnDefs = list(list(className = 'dt-center', targets = 2:3))),
#             caption = "Missing Values in Each Variable",
#             colnames = c("Variable", "Number of Missing", "Percentage of Missing"))
```


### Vaccination Data
Dataset used in the post was sourced from <a href=https://www.ecdc.europa.eu/en/publications-data/data-covid-19-vaccination-eu-eea> European Centre for Disease Prevention and Control</a>, it provides open source data for infectious diseases. This dataset contains vaccination data for `r length(unique(vacc_se$ReportingCountry))` European countries, there are `r ncol(vacc_se)` variables and `r nrow(vacc_se)` observations and all data are up to date, data dictionary is shown in Figure \@ref(fig:vacc).

```{r vacc, fig.cap = "Data Dictionary"}
tribble(
  ~Variable, ~Class, ~Description,
  #---|----|--------------
  "YearWeekISO", "Character", "Date when the vaccine was administered. Only weeks are allowed (e.g. “2021-W01”).",
  "FirstDose", "Numeric", "Number of first dose vaccine administered to individuals during the reporting week.",
  "FirstDoseRefused", "Numeric", "Number of individuals refusing the first vaccine dose.",
  "SeconDose", "Numeric", "Number of second dose vaccine administered to individuals during the reporting week.",
  "Unknown Dose", "Numeric", "Number of doses administered during the reporting week where the type of dose (first or second) was not specified.",
  "NumberDosesReceived", "Numeric", "Number of vaccine doses distributed by the manufacturers to the country during the reporting week.",
  "Region", "Character", "Country code",
  "Polulation", "Numeric", "Age-specific population for the country",
  "Reporting Country", "Character", "Two-letter county code",
  "Target Group","Character", "Target group for vaccination. “ALL” for the overall figures, “HCW” for healthcare workers and age-groups.",
  "Vaccine", "Charater", "Name of vaccine. AZ = AstraZeneca, BECNBG (previously CN) = Beijing CNBG, BHACOV = Bharat, COM = Pfizer/BioNTech, JANSS = Janssen, HAYATVAC = Hayat VAC, MOD = Moderna, QAZVAQ = QazCovid-In, SIICOV = Covishield – SII, SIN = Sinovac, SPU = Sputnik V - Gamaleya, SRCVB = EpiVacCorona – SRCVB, WUCNBG = Inactivated – Wuhan CNBG, UNK = UNKNOWN, ZFUZ = Sino-Uzbek - ZF-UZVAC",
  "Denominator", "Numeric", "Population denominators for target groups.") %>% 
    datatable(options = list(pageLength = 5,
                            columnDefs = list(list(className = 'dt-left', targets = 1:3))),
            caption = "Data Dictionary")
```
This post mainly focuses on analysing the data for Sweden, Figure \@ref(fig:missing) shows the number of missing values and their percentage in each variable in the sub-dataset of Sweden. Variable FirstDoseRefused and Denominator will not be used for the analysis and they can be omitted. As the number of vaccine doses distributed to the country were not on a weekly basis, therefore there are 94% of missing values.

```{r missing, fig.cap = "Summary of Missing Value in Sub-dataset for Sweden"}
# vacc_se %>% 
#   filter(ReportingCountry == "SE") %>% 
# miss_var_summary() %>% 
#     datatable(options = list(pageLength = 5,
#                             columnDefs = list(list(className = 'dt-center', targets = 2:3))),
#             caption = "Missing Values in Each Variable",
#             colnames = c("Variable", "Number of Missing", "Percentage of Missing"))
```

## Data Exploration 🔍

### COVID-19 Survey
A survey has been done to examine the impact of COVID-19 on people's mental health and alcohol consumption in 21 European countries between April and July 2020 after COVID-19 hit Europe. Figure \@ref(fig:mental) shows the response on how they find the current situation of the spread of COVID-19  of the 808 Swedes who participated on the survey. They were divided into 3 age groups, 18-34 years, 35-54 years and 55-98 years. As it can be seen from the graph, approximately 60% of the people from each group found the pandemic stressful, and this is not surprising. You would probably think that this may cause an increase in alcohol consumption as people would turn to alcohol to cope with the stress, however, Figure \@ref(fig:alco) reveals a different story. Compare to some of the European countries like United Kingdom, Ireland, Denmark, the Swedes seemed to consume less alcohol during the pandemic, with the majority of the people did not change their drinking habits and people who drank less were in the second largest group, only a small proportion of people consuming more alcohol. This pattern was more obvious in the younger cohorts, there was a large proportion of people who is age between 18-24 years consume less more alcohol than the other age groups (see Figure \@ref(fig:alco-age)). The restrictions implemented by the Swedish government have greatly reduced social gatherings where alcohol consumption usually was expected and often pressured to happen, and this is the reason for the reduction of alcohol use among the younger cohorts as they generally consume more alcohol in social events (Movendi, 2020). A research found that during March - April 2020, alcohol consumption among Swedish adults and youth in the lowest in the EU (Nordic Alcohol and Drug Policy Network, n.d). In contrast, there were quite a lot of people in the other 3 countries consumed more alcohol than pre-pandemic, especially for United Kingdom and Ireland, the top answer was "slightly more" and it was the second most answer for Denmark. One possible explanation for this diverging patterns could be the less restrict alcohol policies in the United Kingdom, the U.K government declares alcohol shops are essential businesses (Forbes, 2020), whereas in Sweden, there are more restrictive policies in place, such as, alcohol can only be purchased from a government owned liquor store and alcohol ban after 10pm etc.

```{r surv-cleaning}
surv_long <- surv_se %>% 
  select(!c(1:6, "audit_sum", "country", "week", "weight", "popweight")) %>% 
  relocate(change_income, .after = consumer) %>% 
  mutate(diagnose = as.character(diagnose),
         sex = as.character(sex)) %>% 
  pivot_longer(cols = "change_income":"change_illdrugs",
               names_to = "covid_survey",
               values_to = "score")

surv_long_se <- surv_long %>% 
    filter(country_label == "Sweden")
  
```

#### Mental Health

```{r mental, fig.cap = "Mental Health Survey Question"}
surv_long_se$agegr[surv_long_se$agegr == "1"] <- "18-34 years"
surv_long_se$agegr[surv_long_se$agegr == "2"] <- "35-54 years"
surv_long_se$agegr[surv_long_se$agegr == "3"] <- "55-98 years"

n1 <- surv_se %>% 
  filter(country_label == "Sweden",
         agegr == "1") %>% 
  count()

n2 <- surv_se %>% 
  filter(country_label == "Sweden",
         agegr == "2") %>% 
  count()

n3 <- n2 <- surv_se %>% 
  filter(country_label == "Sweden",
         agegr == "3") %>% 
  count()

newcol <- c(0.67, 0.31, 0.68, 0.32, 0.64, 0.36)

newcol <- percent(newcol, 1)

df <- surv_long_se %>% 
  filter(covid_survey == "distress") %>% 
  group_by(agegr, score) %>% 
  summarise(count = n()) 

df$pct <- newcol

df %>% 
  ggplot(aes(x = agegr,
             y = count,
             fill = score)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_manual(values = c("#fc8d62", "#66c2a5"),
                    labels = c("Yes", "No")) +
  scale_y_continuous(labels = scales::percent) +
  geom_text(aes(label = pct), 
            position = position_fill(vjust = 0.5),
            size = 3,
            colour = "white") +
  theme_light() +
  labs(x = "Age Group",
       y = "Percentage",
       fill = "Answer",
       title = "Have you found the spread of COVID-19 stressful?",
       subtitle = "Year: 2020",
       caption = "Data Source: Swedish COVID-19 Data Portal") +
  theme(plot.caption.position = "panel",
        plot.caption = element_text(hjust = 0))

# ggplotly(plot1) %>% 
#     layout(title = list(text = paste0("Have you found the spread of COVID-19 stressful?",
#                                     "<br>",
#                                     "<sup>",
#                                     "Year: 2020",
#                                     "</sup>")),
#            annotations = list(x = 1, y = -5.4, 
#                               text = "Data Source: Swedish COVID-19 Data Portal",
#                               xref='paper',
#                               showarrow = F,
#                               font=list(size=6)))

```

```{r finance}
### Financial Status
# surv_long_se$residence[surv_long_se$residence == "1"] <- "Village/fram"
# surv_long_se$residence[surv_long_se$residence == "2"] <- "Small city"
# surv_long_se$residence[surv_long_se$residence == "3"] <- "Medium-size city"
# surv_long_se$residence[surv_long_se$residence == "4"] <- "Large city"
# surv_long_se$residence[surv_long_se$residence == "5"] <- "Very large city"
# 
# 
# surv_long_se %>%
#   filter(covid_survey == "change_financial",
#          residence != "." ) %>%
#   group_by(residence, score) %>%
#   summarise(count = n()) %>%
#   mutate(score = reorder_within(score, count, residence)) %>%
#   ungroup() %>%
#   ggplot(aes(x = score,
#              y = count,
#              fill = residence)) +
#   geom_col() +
#   facet_wrap(vars(residence),
#              scales = "free") +
#   scale_x_reordered() +
#   scale_fill_brewer(palette = "Set2") +
#   geom_text(aes(label = count, fill = NULL, vjust = 1.2),
#             size = 2.5,
#             colour = "white") +
#   theme_light() +
#   theme(legend.position = "none") +
#   labs(x = "Answer",
#        y = "Total Count",
#        title = "In the past month, have you experienced any negative consequences \nto your employment or financial situation due to the spread of COVID-19?",
#        fill = "Answer",
#        subtitle = "Year: 2020",
#        caption = "Answer: 0 = not at all, 1 = to some degree, 2 = to a substantial degree, 3 = to a very high degree \nData Source: Swedish COVID-19 Data Portal") +
#   theme(plot.caption.position = "panel",
#         plot.caption = element_text(hjust = 0))

```
#### Drinking Behaviour

```{r alco, fig.cap = "Alcohol Consumption Survey Question"}
surv_long %>% 
  filter(country_label %in% c("United Kingdom",
                              "Denmark",
                              "Sweden",
                              "Ireland"),
         covid_survey == "change_alc_freq",
         score != "." ) %>% 
  group_by(country_label, score) %>% 
  summarise(count = n()) %>% 
  mutate(score = reorder_within(score, count, country_label)) %>% 
  ungroup() %>% 
  ggplot(aes(x = score,
             y = count,
             fill = country_label)) +
  geom_col() +
  facet_wrap(vars(country_label),
             scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  scale_fill_brewer(palette = "Set2") +
  geom_text(aes(label = count, fill = NULL, hjust = 1),
            size = 3,
            colour = "white") +
  theme_light() +
  labs(x = "Survey Answer",
       y = "Total Count",
       title = "Did you drink alcohol less or more often in the past month?",
       subtitle = "Year: 2020",
       caption = "Answer: 1 = much less, 2 = slightly less, 3 = no change, 4 = slightly more, 5 = much more \nData Source: Swedish COVID-19 Data Portal") +
  theme(plot.caption.position = "panel",
        plot.caption = element_text(hjust = 0),
        legend.position = "none",
        axis.text.y = element_text(size = 8))

```
```{r alco-age, fig.cap="Alcohol Consumption in Sweden, by Age"}

surv_long_se %>% 
  filter(covid_survey == "change_alc_freq",
         score != "." ) %>% 
  group_by(agegr, score) %>% 
  summarise(count = n()) %>% 
  mutate(score = reorder_within(score, count, agegr)) %>% 
  ggplot(aes(x = score,
             y = count,
             fill = agegr)) +
  geom_col() +
  facet_wrap(vars(agegr),
             scales = "free") +
  scale_x_reordered() +
  scale_fill_brewer(palette = "Set2") +
  geom_text(aes(label = count, fill = NULL, vjust = -0.3),
            size = 3,
            colour = "black") +
  theme_light() +
  labs(x = "Survey Answer",
       y = "Total Count",
       title = "Did you drink alcohol less or more often in the past month?",
       subtitle = "Year: 2020",
       caption = "Answer: 1 = much less, 2 = slightly less, 3 = no change, 4 = slightly more, 5 = much more \nData Source: Swedish COVID-19 Data Portal") +
  theme(plot.caption.position = "panel",
        plot.caption = element_text(hjust = 0),
        legend.position = "none",
        axis.text.y = element_text(size = 8))
```

```{r smoke}
# smoke
# p2 <- surv_long_se %>% 
#   filter(covid_survey == "change_smoke",
#          score != "." ) %>% 
#   group_by(score) %>% 
#   summarise(count = n()) %>% 
#   ggplot(aes(x = fct_reorder(score, count),
#              y = count,
#              fill = score)) +
#   geom_col() +
#   coord_flip() +
#   scale_fill_manual(values = c("#fc8d62","#fc8d62", "#fc8d62", "#fc8d62", "#fc8d62", "#fc8d62")) +
#   scale_x_discrete(labels = c("much more", "slightly less", "slightly more", "much less", "no change", "do not consume")) +
#   scale_y_continuous(breaks=seq(0, 700, 100)) +
#   geom_text(aes(label = count, fill = NULL, hjust = -0.1),
#             size = 3,
#             colour = "#fc8d62",
#             angel = 90) +
#   theme_light() +
#   theme(legend.position = "none",
#         axis.text.y = element_text(size = 8)) +
#   labs(x = "Answer",
#        y = "Total Count",
#        title = "Did you smoke less or more often in the past month?",
#        subtitle = "Year: 2020",
#        caption = "Data Source: Swedish COVID-19 Data Portal") +
#   theme(plot.caption.position = "plot",
#         plot.caption = element_text(hjust = 0))
# 
# 
# (p1 / p2 ) + plot_annotation(tag_levels = 'A')
```


```{r drugs}
# # cannabis
# p3 <- surv_long_se %>% 
#   filter(covid_survey == "change_cannabis",
#          score != "." ) %>% 
#   group_by(score) %>% 
#   summarise(count = n()) %>% 
#   ggplot(aes(x = fct_reorder(score, count),
#              y = count,
#              fill = score)) +
#   geom_col() +
#   coord_flip() +
#   scale_fill_manual(values = c("#8da0cb","#8da0cb", "#8da0cb", "#8da0cb", "#8da0cb", "#8da0cb")) +
#   scale_x_discrete(labels = c("slightly less", "much more", "much less", "slightly more", "no change", "do not consume")) +
#   geom_text(aes(label = count, fill = NULL, hjust = -0.1),
#             size = 3,
#             colour = "#8da0cb") +
#   theme_light() +
#   labs(x = "Answer",
#        y = "Total Count",
#        title = "Did you consume cannabis less or more often in the past month?",
#        subtitle = "Year: 2020") +
#   theme(plot.caption.position = "panel",
#         plot.caption = element_text(hjust = 0),
#         legend.position = "none",
#         axis.text.y = element_text(size = 8))
# 
# 
# # drugs
# p4 <- surv_long_se %>% 
#   filter(covid_survey == "change_illdrugs",
#          score != "." ) %>% 
#   group_by(score) %>% 
#   summarise(count = n()) %>% 
#   ggplot(aes(x = fct_reorder(score, count),
#              y = count,
#              fill = score)) +
#   geom_col() +
#   coord_flip() +
#   scale_fill_manual(values = c("#e78ac3","#e78ac3", "#e78ac3", "#e78ac3", "#e78ac3", "#e78ac3")) +
#   scale_x_discrete(labels = c("slightly more", "much less", "no change", "do not consume")) +
#   geom_text(aes(label = count, fill = NULL, hjust = -0.1),
#             size = 3,
#             colour = "#e78ac3") +
#   theme_light() +
#   labs(x = "Answer",
#        y = "Total Count",
#        title = "Did you use illegal drugs less or more often in the past month?",
#        subtitle = "Year: 2020",
#        caption = "Data Source: Swedish COVID-19 Data Portal") +
#   theme(plot.caption.position = "panel",
#         plot.caption = element_text(hjust = 0),
#         legend.position = "none",
#         axis.text.y = element_text(size = 8))
# 
# (p3 / p4) + plot_annotation(tag_levels = 'A')
```


### Current COVID-19 Situation in Sweden

#### New Confirmed Cases and Deaths

Sweden has recorded three waves sofar, the two major ones are in the late 2020 to early 2021 and March to June this year and the number of confirmed cases per million people were the highest among its neighbouring countries, Germany, Norway, Denmark, Finland, as shown in the Figure \@ref(fig:newcases). The Swedish government's soft touch approach on dealing with the virus among the European countries has drawn worldwide attention and criticism (ABC news, 2020), and when the third wave hit, Sweden had already opted against lockdowns, instead relying on mostly voluntary measures (Reuters, 2021), the confirm cases recorded in April and May surpassed most of the European countries. From Figure \@ref(fig:newcases), we can also see that, there is more likely to be another wave in the coming months, but the increase in confirmed cases is not yet significant, let's hope the vaccination can be of assistance in flatting the curve. The daily number of deaths per million people seems relatively low in the graph, however, if we take the daily numbers and calculate the cumulative daily number of deaths, Sweden again, is well above its neighbours, see Figure \@ref(fig:newdeaths). For a country that only have 10 million population, only one-eight of the population in Germany, this is very concerning.

<style>
.html-widget {
    margin: auto;
}
</style>

```{r}
se <- world %>% 
  filter(location == "Sweden") %>% 
  select(location,
         date,
         new_cases_per_million,
         new_deaths_per_million,
         total_vaccinations,
         new_vaccinations,
         new_vaccinations_smoothed,
         people_vaccinated,
         people_vaccinated_per_hundred,
         people_fully_vaccinated,
         people_fully_vaccinated_per_hundred)
```

```{r newcases, fig.cap="Daily Confirmed Cases per Million People"}

deaths_se <- se %>% 
select(date, new_deaths_per_million)


plot1 <- world %>% 
  select(location,
         date,
         new_cases_per_million) %>% 
  filter(location %in% c("Sweden",
                         "Germany",
                         "Denmark",
                         "Finland",
                         "Norway")) %>% 
  ggplot(aes(x = date,
             y = new_cases_per_million,
             colour = location)) +
  geom_line(size = 0.3) +
  geom_point(size = 0.2) +
  geom_point(data = deaths_se,
             aes(x = date,
                 y = new_deaths_per_million,
                 colour = "Deaths in Sweden"),
             size = 0.4) +
  scale_colour_brewer(palette = "Set2") +
  theme_light() +
  labs(x = "Date",
       y = "Daily Confirmed New Cases",
       colour = "Country",
       title = "Daily New Cases per Million People",
       subtitle = "Year: Feb 2020 - Sep 2021") +
  scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = scales::comma) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.text.y = element_text(size = 8))

ggplotly(plot1) %>%
  layout(title = list(text = paste0("Daily New Cases per Million People",
                                    "<br>",
                                    "<sup>",
                                    "Year: Feb 2020 - Sep 2021",
                                    "</sup>")),
         annotations = list(x = 1, y = -0.5,
                              text = "Data Source: Our World Data",
                              showarrow = F,
                              font=list(size=8)))
  
```


```{r newdeaths, fig.cap="Accumulative Daily Deaths"}
plot2 <- world %>% 
  select(location,
         date,
         new_deaths_per_million) %>% 
  filter(location %in% c("Sweden",
                         "Germany",
                         "Denmark",
                         "Finland",
                         "Norway"),
         !is.na(new_deaths_per_million)) %>% 
  group_by(location) %>% 
  mutate(cumulative_sum = cumsum(new_deaths_per_million)) %>% 
  ggplot(aes(x = date,
             y = cumulative_sum,
             colour = location)) +
  geom_line(size = 0.3) +
  geom_point(size = 0.2) +
  scale_colour_brewer(palette = "Set2") +
  theme_light() +
  labs(x = "Date",
       y = "Number of Deaths",
       colour = "Country",
       title = "Accumulative Deaths per Million People",
       subtitle = "Year: Feb 2020 - Sep 2021") +
  scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = scales::comma) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.text.y = element_text(size = 8))

ggplotly(plot2) %>%
  layout(title = list(text = paste0("Accumulative Deaths per Million People",
                                    "<br>",
                                    "<sup>",
                                    "Year: Feb 2020 - Sep 2021",
                                    "</sup>")),
         annotations = list(x = 1, y = -0.5,
                              text = "Data Source: Our World Data",
                              showarrow = F,
                              font=list(size=8)))
```

#### Vaccination

After a year since the coronavirus emerged, the first vaccines developed by the global effort started being distributed round the world in the late 2020. Sweden began their vaccine programme on the 27th Dec after the approval of Pfizer vaccine. By September this year, Sweden has reached approximately 60% of the population fully vaccinated and almost 70% that had one dose, as shown in Figure \@ref(fig:vacc-se), total of 6,131,967 people out of 10 million population are fully vaccinated and the numbers are still increasing. Despite the fact that the Swedish COVID approach was not the best, but their vaccine roll out is a winner. The deaths rate for the third wave was much lower than the second wave with lockdowns shunned and masks banned. Sweden's vaccine roll out consists of 4 phases, phase 1 is for people who live in care homes for elderly, who have at-home care and who live with home, phase 2 is for people aged 65-69, phase 3 is for people aged 60-64 and phase 4 is for other adults aged 18-59, oldest first (The Local, 2021). Figure \@ref(ref:vacc-age) shows the timeline of fully vaccinate people in each age group and how the roll out imposed, the seniors groups were vaccinated first followed by the younger groups.

```{r}
# remove "w" in string
vacc_se$YearWeekISO <- gsub("[[:alpha:]]", "", vacc_se$YearWeekISO)

# add month column
vacc_se$month <- strsplit(vacc_se$YearWeekISO, "-")

vacc_se$month <- sapply(vacc_se$month,function(x) {
  year_week <- unlist(x)
  year <- year_week[1]
  week <- year_week[2]
  start_date <- as.Date(paste0(year,'-01-01'))
  date <- start_date+weeks(week)
  return (as.character(date))
})

vacc_se$month <- month(vacc_se$month)

vacc_se$month <- ifelse(str_starts(vacc_se$YearWeekISO, "2020"), "12", vacc_se$month)

vacc_se$date <- paste(vacc_se$YearWeekISO, vacc_se$month, sep = "/")

# vacc_se$year <- ifelse(str_starts(vacc_se$month, "12"), "2020", "2021")
# 
# vacc_se$date <- paste(vacc_se$year, vacc_se$month, sep = "-")

# vacc_se <- vacc_se %>%
#   separate(YearWeekISO, c("year", "week"), "-") %>%
#   relocate(month, .after = week)
# 
# vacc_se$yearMonth <- paste(vacc_se$year, vacc_se$month, sep = "-")

```


```{r vacc-se, fig.cap="Accumulative Vaccines Administered"}
plot3 <- world %>% 
  select(location,
         date,
         people_vaccinated,
         people_fully_vaccinated) %>% 
  rename("Fully Vaccinated" = people_fully_vaccinated,
         "First Dose" = people_vaccinated) %>% 
  filter(location == "Sweden") %>% 
  pivot_longer(cols = c("First Dose", 
                        "Fully Vaccinated"),
               names_to = "dose",
               values_to = "entries") %>% 
  ggplot(aes(x = date,
             y = entries,
             group = dose,
             fill = dose)) +
  geom_area(alpha = 0.8, size = 0.5, colour = "white") +
  scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(values = c("#66c2a5", "#fc8d62")) +
  theme_light() +
  labs(x = "Date",
       y = "Number of Doses Administered",
       title = "Accumulative Vaccines Administered, by Month",
       fill = "Dose",
       caption = "European Centre for Disease Prevention and Control") +
  theme(plot.caption.position = "panel",
        plot.caption = element_text(hjust = 0),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.text.y = element_text(size = 8))

ggplotly(plot3) %>% 
       layout(title = list(text = paste0("Accumulative Vaccines Administered, by Month",
                                    "<br>",
                                    "<sup>",
                                    "Year: Dec 2020 - Sep 2021",
                                    "</sup>")),
           annotations = list(x = 1, y = -0.5,
                              text = "Data Source: European Centre for Disease Prevention and Control",
                              showarrow = F,
                              font=list(size=8)))

```


```{r vacc-age, fig.cap="Weekly Accumulative Number of Fully Vaccinated People"}

plot4 <- vacc_se %>% 
  select(date, 
         ReportingCountry,
         SecondDose,
         TargetGroup) %>% 
  filter(ReportingCountry == "SE",
         TargetGroup %in% c("Age18_24", 
                            "Age25_49", 
                            "Age50_59", 
                            "Age60_69", 
                            "Age70_79", 
                            "Age80+")) %>% 
  group_by(TargetGroup, date) %>%
  summarise(count = sum(SecondDose)) %>%
  mutate(cumulative_sum = cumsum(count)) %>% 
  ungroup() %>% 
  ggplot(aes(x = date,
             y = cumulative_sum,
             group = TargetGroup,
             colour = TargetGroup)) +
  geom_line() +
  geom_point(size = 0.6) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_brewer(palette = "Set2",
                      labels = c("Age 18-24",
                                 "Age 25-49",
                                 "Age 50-59",
                                 "Age 60-69",
                                 "Age 70-79",
                                 "Age 80+")) +
  theme_light() +
  labs(x = "Date",
       y = "Number of Doses Administered",
       title = "Weekly Accumulative Number of Fully Vaccinated People, by Age",
       colour = "Age Group",
       caption = "Source: European Centre for Disease Prevention and Control") +
  theme(plot.caption.position = "panel",
        plot.caption = element_text(hjust = 0),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.text.y = element_text(size = 8)) 

ggplotly(plot4) %>%
       layout(title = list(text = paste0("Weekly Accumulative Number of Fully Vaccinated People, by Age",
                                    "<br>",
                                    "<sup>",
                                    "Year: Dec 2020 - Sep 2021, date is year-week/month",
                                    "</sup>")))

```


```{r}
# vacc_se$Vaccine[vacc_se$Vaccine == "COM"] <- "Pfizer"
# vacc_se$Vaccine[vacc_se$Vaccine == "MOD"] <- "Moderna"
# vacc_se$Vaccine[vacc_se$Vaccine == "AZ"] <- "AstraZenecar"
# vacc_se$Vaccine[vacc_se$Vaccine == "UNK"] <- "Unknown"
# 
# 
# vacc_se %>% 
#   filter(ReportingCountry == "SE",
#          TargetGroup == "ALL") %>% 
#   group_by(Vaccine) %>% 
#   summarise(count = sum(SecondDose)) %>% 
#   ggplot(aes(area = count,
#              fill = Vaccine,
#              label = paste0(Vaccine, "\n", prettyNum(count, big.mark = ",")),
#              subgroup = Vaccine)) +
#   geom_treemap() +
#   geom_treemap_subgroup_border(colour = "white", size = 1) +
#   geom_treemap_text(color = "white", size = 18, min.size = 1) +
#   scale_fill_brewer(palette = "Set2") +
#   labs(fill = "Vaccine",
#        title = "Type of Vaccines Adminstered in Sweden",
#        subtitle = "Year: Dec 2020 - Sep 2021",
#        caption = "Data Source: Swedish COVID-19 Data Portal") +
#   theme(axis.text.y = element_text(size = 6),
#         plot.caption.position = "panel",
#         plot.caption = element_text(hjust = 0)) 

```







## Reference

ECDC, European Centre for Disease Prevention and Control. (2021, September, 9). "Data on COVID-19 vaccination in the EU/EEA", <https://www.ecdc.europa.eu/en/publications-data/data-covid-19-vaccination-eu-eea>

Mathieu, E., Ritchie, H., Ortiz-Ospina, E. et al. A global database of COVID-19 vaccinations. Nat Hum Behav (2021)

Movendi. (2021, May, 26). "Sweden: Youth Consumes Less Alcohol During Coronavirus Crisis", <https://movendi.ngo/news/2020/05/26/sweden-youth-consumes-less-alcohol-during-coronavirus-crisis/>

Forbes. (2020, Mar, 25). "Coronavirus: U.K Government Declares Alcohol Shops ‘Essential Businesses", <https://www.forbes.com/sites/felipeschrieberg/2020/03/25/coronavirus-uk-government-declares-alcohol-shops-essential-businesses/?sh=434235ea4b1c>

Nordic Alcohol and Drug Policy Network. (2020, n.d). "Sweden - Consumption trends", <https://www.nordicalcohol.org/sweden-consumption-trends>

ABC news. (2020, Nove, 14). "Sweden maintains controversial COVID-19 strategy despite surge in coronavirus cases", <https://www.abc.net.au/news/2020-11-14/sweden-second-wave-covid-19-strategy-surge-cases/12884194>

Reuters. (2021, May). "Third COVID-19 wave is ebbing fast in Sweden, expert says", <https://nationalpost.com/pmn/health-pmn/third-covid-19-wave-is-ebbing-fast-in-sweden-expert-says>

The Local. (2021, April 14). "EXPLAINED: When will I get my Covid-19 vaccination in Sweden?", <https://www.thelocal.se/20210414/explained-when-will-i-get-my-covid-19-vaccination-in-sweden/>