---
title: "Covid in Switzerland 🇨🇭"
description: |
  Total Number of Confirmed COVID-19 Cases, Deaths, Tests, Hospitalisations and Vaccines Administered in Switzerland.
preview: thumbnail_1.png
author:
  - name: Jaffie Yang
    url: https://example.com/norajones
date: 08-29-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      out.width = "80%",
                      fig.align = "center")
```

```{r packages, message = FALSE}
library(here)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(lubridate)
library(naniar)
library(kableExtra)
library(formattable)
library(DT)
library(gt)
library(sf)
library(rworldmap)
library(plotly)
library(leaflet)
library(rjson)
library(geojsonio)
library(tidycovid19)
library(zoo)
```

```{r read-data}
cases <- read_csv(here("data/COVID19IntCases.csv"))
death <- read_csv(here("data/COVID19death.csv"))
vacc <- read_csv(here("data/COVID19vaccine.csv"))
hosp <- read_csv(here("data/COVID19hosp.csv"))
test <- read_csv(here("data/COVID19test.csv"))
#poly <- geojson_read(here("data/countries.geojson", what = "sp"))

```


```{r join-data}
sc <- cases %>% 
  filter(geoRegionName == "Switzerland") %>% 
  select(geoRegion:entries) %>% 
  rename(confirmed_cases = "entries")

sd <- death %>% 
  filter(geoRegion == "CH") %>% 
  select(datum:entries) %>% 
  rename(total_deaths = "entries",
         date = "datum")

sh <- hosp %>% 
  filter(geoRegion == "CH") %>% 
  select(datum:entries, "pop") %>% 
  rename(hospitalisation = "entries",
         date = "datum") 

st <- test %>% 
  filter(geoRegion == "CH") %>% 
  select(datum:"entries_neg") %>% 
  rename(total_test = "entries",
         total_pos_test = "entries_pos",
         total_neg_test = "entries_neg",
         date = "datum")

st <- st[32:581, ]


s_vac <- vacc %>% 
  filter(geoRegion == "CH") %>% 
  select(date:entries) %>% 
  rename(total_vac_admin = "entries")

s_cov <- sc %>% 
  left_join(st,
            by = "date") %>% 
  left_join(sd,
            by = "date") %>% 
  left_join(sh,
            by = "date") 
```

## Data Description 📄

The five separate data sets used in the post were sourced from opendata.swiss, it is the Swiss public administration's central portal for open government data (opendata.swiss, 2021). Four of the five data sets were joined together to create a new data set call s_cov, it contains `r ncol(s_cov)` variables and `r nrow(s_cov)` observations. Data dictionary and number and percentage of missings in each variable are as follows:

|**Variable**      |**Class** |**Description**                    |
|:-----------------|:---------|:----------------------------------|
|geoRegion         |character |Country geocode                    |
|geoRegionName     |character |Country name                       |
|date              |date      |Date                               |
|daily_confirmed   |numeric   |Daily confirmed Covid-19 cases     | 
|daily_test        |numeric   |Daily total test numbers           |
|daily_pos_test    |numeric   |Daily total positive test          |
|daily_neg_test    |numeric   |Daily total negative test          |
|daily_total_deaths|numeric   |Daily total deaths                 | 
|daily_hosp        |numeric   |Daily total hospitalisation numbers|


```{r}
miss_var_summary(s_cov) %>% 
  datatable(options = list(pageLength = 4))
```

<br><br>

The other data set is call s_vac, it has `r ncol(s_vac)` variables and `r nrow(s_vac)` observations. Tables of Data dictionary and missing values in each variables are as follows:

|**Variable**      |**Class** |**Description**                                  |
|:-----------------|:---------|:------------------------------------------------|
|date              |date      |Date                                             |
|geoRegion         |character |Country geocode                                  |
|vaccines          |date      |Type of vaccines administered: Moderna or Pfizer |
|daily_confirmed   |numeric   |Daily confirmed Covid-19 cases                   |   

```{r}
miss_var_summary(s_vac) %>% 
  datatable()
```

<br><br>

## Data Story 📖

### Total Confirmed COVID-19 Cases in Europe

```{r map-data} 
# EU contries
eu <- cases %>% 
  filter(geoRegionName %in% c("Albania", "Andorra", "Austria", "Belarus", "Belgium", "Bosnia and Herzegovina", "Bulgaria",
                          "Croatia", "Czech Republic", "Denmark", "Estonia", "Finland", "France", "Germany", "Greece",
                          "Holy See", "Hungary", "Iceland", "Ireland", "Italy", "Latvia", "Liechtenstein", "Lithuania",
                          "Luxembourg", "Malta", "Moldova", "Monaco", "Montenegro", "Netherlands", "North Macedonia",
                          "Norway", "Poland", "Portugal", "Romania", "Russia", "San Marino", "Serbia", "Slovakia",
                          "Slovenia", "Spain", "Sweden", "Switzerland", "Ukraine", "United Kingdom")) %>% 
  group_by(geoRegionName) %>% 
  summarise(count = sum(entries)) 

# get world map
worldMap <- getMap()

# Select only the index of selected EU countries
indEU <- which(worldMap$NAME %in% eu$geoRegionName)

# Extract longitude and latitude border's coordinates  
europeCoords <- lapply(indEU, function(i){
  df <- data.frame(worldMap@polygons[[i]]@Polygons[[1]]@coords)
  df$region =as.character(worldMap$NAME[i])
  colnames(df) <- list("long", "lat", "region")
  return(df)
})

europeCoords <- do.call("rbind", europeCoords)

#join data
eu_map <- eu %>% 
  rename(region = "geoRegionName") %>% 
  right_join(europeCoords,
             by = "region")
```


As Figure \@ref(fig:map) shown below, Russian, Franch and the United Kingdom have the most confirmed cases, followed by Italy, Germany and Spain. Switzerland are amongst countries that have less confirmed cases.

<style>
.html-widget {
    margin: auto;
}
</style>

```{r map, fig.cap = "Total Confirmed Covid-19 Cases in Europe by 26/08/2021"}
map <- ggplot() + 
  geom_polygon(data = eu_map, 
               aes(x = long, 
                   y = lat, 
                   group = region, 
                   fill = count),
               colour = "black", size = 0.1) +
  coord_map(xlim = c(-15, 40),  
            ylim = c(30, 80))+ 
  scale_fill_gradient(n.breaks = 6,
                      name = "Total Confirmed Cases", 
                      low = "#bffae6", 
                      high = "#185a9d", 
                      na.value = "grey50") +
  theme(#panel.grid.minor = element_line(colour = NA), panel.grid.minor = element_line(colour = NA),
               panel.background = element_rect(fill = "#fffbf0"),
               plot.background= element_rect(fill = "#fffbf0", colour = "#fffbf0"),
               legend.background = element_rect(fill = "#fffbf0", colour = "#fffbf0"),
               axis.text.x = element_blank(),
               axis.text.y = element_blank(), 
               axis.ticks.x = element_blank(),
               axis.ticks.y = element_blank(), 
               axis.title = element_blank()) +
               #rect = element_blank(),
               #plot.margin = unit(0 * c(-1.5, -1.5, -1.5, -1.5), "lines"))
  labs(title = "Total Confirmed Covid-19 Cases in Europe by 26/08/2021")

 ggplotly(map)
```

```{r covid, fig.cap = "COVID-19 Figures from Feb 2020 to August 2021"}
# 2020
s_cov_1 <- s_cov %>% 
  mutate(Year = year(date),
         Month = month(date)) %>% 
  filter(geoRegionName == "Switzerland",
         Year == "2020") %>% 
  group_by(Month, Year) %>% 
  summarise(across(c("confirmed_cases", "total_test", 
                     "total_deaths", "hospitalisation"), 
                   ~ sum(.x, na.rm = TRUE)))
# 2021
s_cov_2 <- s_cov %>% 
  mutate(Year = year(date),
         Month = month(date)) %>% 
  filter(geoRegionName == "Switzerland",
         Year == "2021") %>% 
  group_by(Month, Year) %>% 
  summarise(across(c("confirmed_cases", "total_test", 
                     "total_deaths", "hospitalisation"), 
                   ~ sum(.x, na.rm = TRUE)))
# merge two tables
s_cov_joined <- rbind(s_cov_1, s_cov_2)

# re-arrange columns
s_cov_joined<- s_cov_joined[, c(2,1, 3, 4, 5, 6)]

# DT table
s_cov_joined%>% 
  datatable(colnames = c("Year", "Month", "Confirmed Cases", "Test Numbers", 
                         "Total Deaths", "Hospitalisation Numbers"),
            options = list(pageLength = 10,
                           columnDefs = list(list(className = 'dt-center', targets = 0:5))),
            extensions = 'KeyTable', 
            class = "cell-border stripe",
            caption = "COVID-19 Figures by Montah and Year",
            filter = "top",
            rownames = FALSE) %>% 
  formatStyle("confirmed_cases",
              background = styleColorBar(s_cov_joined$confirmed_cases, "#b5eaea"),
              backgroundSize = "95% 60%",
              backgroundRepeat = "no-repeat",
              backgroundPosition = "right") %>% 
  formatStyle("total_test",
              background = styleColorBar(s_cov_joined$total_test, "#e4efe7"),
              backgroundSize = "95% 50%",
              backgroundRepeat = "no-repeat",
              backgroundPosition = "right") %>% 
  formatStyle("total_deaths",
              background = styleColorBar(s_cov_joined$total_deaths, "#f0d9ff"),
              backgroundSize = "95% 50%",
              backgroundRepeat = "no-repeat",
              backgroundPosition = "right") %>% 
  formatStyle("hospitalisation",
              background = styleColorBar(s_cov_joined$hospitalisation, "#f6dfeb"),
              backgroundSize = "95% 50%",
              backgroundRepeat = "no-repeat",
              backgroundPosition = "right")

```



```{r eval = FALSE}

# 2020
s_cov_long <- s_cov %>% 
  pivot_longer(cols = daily_confirmed:daily_hosp,
               names_to = "type",
               values_to = "count") %>% 
  mutate(year = as.factor(year(date)),
         month = month(date)) 

s_cov_long %>% 
  filter(type %in% c("daily_confirmed",
                     "daily_test",
                     "daily_total_deaths",
                     "daily_hosp")) %>% 
  group_by(month, type, year) %>% 
  summarise(total_count = sum(count))

  
```


```{r vacc}
# 2020
s_vac_1 <- s_vac %>% 
  pivot_wider(names_from = vaccine,
              values_from = total_vac_admin) %>% 
  mutate(Year = year(date),
         Month = month(date)) %>% 
  filter(Year == "2020") %>% 
  group_by(Month, Year) %>% 
  summarise(across(c("moderna", "pfizer_biontech"), ~ sum(.x, na.rm = TRUE)))


# 2021
s_vac_2 <- s_vac %>% 
  pivot_wider(names_from = vaccine,
              values_from = total_vac_admin) %>% 
  mutate(Year = year(date),
         Month = month(date)) %>% 
  filter(Year == "2021") %>% 
  group_by(Month, Year) %>% 
  summarise(across(c("moderna", "pfizer_biontech"), ~ sum(.x, na.rm = TRUE))) 

# join data for 2020 and 2021
s_vac_joined <- rbind(s_vac_1, s_vac_2)

s_vac_joined[, c(2,1, 3, 4)]%>% 
  as.data.frame() %>% 
  formattable(list(area(col = 3:4) ~ color_tile("white", "#D45969")))
  
```






## References

opendata.swiss. (2021, August 27). "COVID-19 Switzerland", <https://opendata.swiss/en/dataset/covid-19-schweiz?